Slavnov products and KP tau functions

### 2025-04-14 draft.tex
add baker-akhiezer function

\section{Miwa coordinates}

@@

It is easy to show that that the tau function~(\ref{eq:tau-function})
admits an integral representation of the form
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) = 
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\bm{\xi}(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \; ,
    \right)
\end{equation}
where 
\begin{equation}
  \bm{\xi}(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p}\; ,
\end{equation}
and the integration contour \(\gamma_k\) encloses all points
\(\{z_j\}_{j=1}^n\) but avoids the Bethe root \(v_k\).  For a detailed
proof of this statement, one may check the
reference~\cite{Araujo:2021ghu}.  The only difference now is that now
we need to be carefull with the integration contour \(\gamma_k\).

One of the advantages of this representation is that now it makes
sense to consider the limit of the infinite chain, \(n\to \infty\).

@@

It seems that we can consider a basis of functions \(\mathcal{Y}_j(z)\) for the Baker-Akhiezer
functions. I think that this object might say something interesting. In particular, we can
explicitly build solutions of the KP equation. In fact, the functions \(\mathcal{Y}_j\)
above have poles at the Bethe roots, it is interesting. 
\begin{itemize}
\item[\faCheck] What is the pole and root structure of the tau and Baker-Akhiezer (BA) functions?
  In particular, what happens at the Bethe roots?
\item[\faCheck] Basis tau functions
\item[\faCheck] Write the cases n=2 and n=3
\item[\faCheck] What the general case for all \(z_k\) approaching to a single
  value \(w\). 
\item[\faCheck] Write these tau functions in terms of Miwa coordinates
\item I need to calculate the Baker-Akhiezer function for this model using the Japanese formula
  in terms of tau functions. The tau functions above is know. I think that this BA function
  will be similar to the soliton-like solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}. 
  My functions seem to have a pole of order \(N\) at infinity. And \(N\) zeros at the Bethe roots. 
\item It turns out that the Baker-Akhiezer function is the appropriate object to discuss
  the Bethe roots. 
  \item I think that these tau functions come from algebro geometric tau functions. 
  \item I can try to study the dynamics closer to the Bethe roots, I think that the
    \(\Omega\)'s will have some interesting expansions, some of them might vanish. 
  \item This might become a Riemann Hilbert problem
  \item Are these tau functions associated to the Calogero-Moser model??
  \item Elliptic case - see slavnov, zabrodin's paper
\end{itemize}

>>>>

\section{Baker-Akhiezer function}

@@

It is easy to show that that the tau function~(\ref{eq:tau-function})
admits an integral representation of the form
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) = 
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \; ,
    \right)
\end{equation}
where 
\begin{equation}
  \xi(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p}\; ,
\end{equation}
and the integration contour \(\gamma_k\) encloses all points
\(\{z_j\}_{j=1}^n\) but avoids the Bethe root \(v_k\).  For a detailed
proof of this statement, one may check the
reference~\cite{Araujo:2021ghu}.  The only difference now is that now
we need to be carefull with the integration contour \(\gamma_k\).
One of the advantages of this representation is that now it makes
sense to consider the limit of the infinite chain, \(n\to \infty\).

For later convencience, let us also define an invertible matrix
\(\mathcal{K}\) by its components
\begin{equation}
    \mathcal{K}_{jk} = 
\oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k}\;,  
\end{equation}
so that the tau function is simply \(\bm{\tau}(\bm{t}, \bm{v}) = \det\mathcal{K}\).

The Baker-Akhiezer funcion is defined in terms of the Japanese formula~\cite{Zabrodin2018}
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \frac{\bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})}{\bm{\tau}(\bm{t}, \bm{v})}
\end{equation}
where 
\begin{equation}
  \bm{t} - [\lambda^{-1}] = 
  \{t_1 - \lambda^{-1}, t_2 - \lambda^{-2}/ 2 ,  t_2 - \lambda^{-3}/3, \dots,  t_p - \lambda^{-p}/p, \dots \}\; .
\end{equation}

Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Let us first write
\begin{subequations}
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t}, \lambda)} & = \exp\left( \sum_{p=1}^\infty t_p \lambda^p \right)
    = \exp\left( \sum_{p=1}^\infty \sum_{j=1}^n \frac{1}{p} z^p \lambda^p \right) \\
    & = \exp\left( \sum_{j=1}^n \sum_{p=1}^\infty  \frac{1}{p} z_j^p \lambda^p \right)
    = \exp\left( - \sum_{j=1}^n \ln( 1 -  z_j \lambda ) \right) \\
    & =  \prod_{j=1}^n \frac{1}{1 -  z_j \lambda} 
  \end{split}
\end{equation}
and we see that these functions have simple poles at \( \lambda = z_j\).

Moreover, 
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
and we can write
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]\\
    & = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \; .
  \end{split}
\end{equation}
\end{subequations}

Using these expressions, we can write 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \det_{j,k}\left[
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \left( 1 - \frac{1}{\lambda w} \right) \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} 
  \right]\; .
\end{equation}
Now, we define the matrix \(\widetilde{\mathcal{K}}(\lambda)\) by its components
\begin{equation}
    \widetilde{\mathcal{K}}_{jk}(\lambda)  = -  \frac{1}{\lambda} 
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{- 1})} \frac{w^{-j -1 }\mathcal{Y}_k(w)}{w - v_k} \; ,
\end{equation}
and we then have 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \bm{\tau}(\bm{t}, \bm{v})
  \det \left( \mathbb{1} +  \mathcal{K}^{-1} \widetilde{\mathcal{K}}(\lambda) \right)\; .
\end{equation}

We finally conclude that the Bakher-Akhiezer function can be written as
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \det \left( \mathbb{1} +  \mathcal{K}^{-1} \widetilde{\mathcal{K}}(\lambda) \right)\; .
\end{equation}

@@

It seems that we can consider a basis of functions \(\mathcal{Y}_j(z)\) for the Baker-Akhiezer
functions. I think that this object might say something interesting. In particular, we can
explicitly build solutions of the KP equation. In fact, the functions \(\mathcal{Y}_j\)
above have poles at the Bethe roots, it is interesting. 
\begin{itemize}
\item[\faCheck] What is the pole and root structure of the tau and Baker-Akhiezer (BA) functions?
  In particular, what happens at the Bethe roots?
\item[\faCheck] Basis tau functions
\item[\faCheck] Write the cases n=2 and n=3
\item[\faCheck] What the general case for all \(z_k\) approaching to a single
  value \(w\). 
\item[\faCheck] Write these tau functions in terms of Miwa coordinates
\item[\faCheck] I need to calculate the Baker-Akhiezer function for this model using the Japanese formula
  in terms of tau functions.
\item Write some explicit expressions for \(N=2\) and \(N=3\). 
\item Better investigate these functions. See Fredholm determinant. 
  I think that this BA function will be similar to the soliton-like solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}. 
  My functions seem to have a pole of order \(N\) at infinity. And \(N\) zeros at the Bethe roots. 
\item It turns out that the Baker-Akhiezer function is the appropriate object to discuss
  the Bethe roots. 
  \item I think that these tau functions come from algebro geometric tau functions. 
  \item I can try to study the dynamics closer to the Bethe roots, I think that the
    \(\Omega\)'s will have some interesting expansions, some of them might vanish. 
  \item This might become a Riemann Hilbert problem
  \item Are these tau functions associated to the Calogero-Moser model??
  \item Elliptic case - see slavnov, zabrodin's paper
\end{itemize}
